Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2020-04-21 Paper/manuscript_v2/manuscript_07_01.tex
Minor edits to sec 5.2

Let us denote $y_t$ as the number of births of day $t$. The observational model is a normal model with parameters the mean function $\mu(t)$ and noise variance $\sigma^2$,
%
\begin{equation*}
y_{t} \sim \Normal(\mu(t),\sigma^2).
\end{equation*}
%
The mean function $\mu(t)$ will be defined as an additive model in the form: 
%
\begin{equation} \label{ch5_eq_mean_brithday}
\mu(t) = f_1(t) + f_2(t) + f_3(t) + f_4(t).
\end{equation}

@@

The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
%
\begin{equation*}
f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
\end{equation*}
%
A horse-shoe prior allows for sparse distributed effects. Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage. Different levels of sparsity can be accommodated by changing the value of $\tau$: with large $\tau$ all the variables have very diffuse priors with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}. 
%More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.

@@

The year effects $f_2(t)$ and week effects $f_3(t)$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper. However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{ch5_sec_periodic}). Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in equation (\ref{ch5_eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in equation (\ref{ch5_eq_q_2}).
We use $J=10$ cosine terms. The lengthscales estimates $\hat{\ell}_2$ and $\hat{\ell}_3$, for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.

>>>>

Let $y_t$ denote the number of births on the $t$'th day. The observation model is a normal distribution with mean function $\mu(t)$ and noise variance $\sigma^2$,
%
\begin{equation*}
y_{t} \sim \Normal(\mu(t),\sigma^2).
\end{equation*}
%
The mean function $\mu(t)$ will be defined as an additive model in the form
%
\begin{equation} \label{ch5_eq_mean_brithday}
\mu(t) = f_1(t) + f_2(t) + f_3(t) + f_4(t).
\end{equation}

@@

The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
%
\begin{equation*}
f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
\end{equation*}
%
A horse-shoe prior allows for sparse distributed effects. Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage. Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}. 
%More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.

@@

The year effects $f_2(t)$ and week effects $f_3(t)$ use a periodic covariance function  and thus do no fit under the main framework of the HSGP approximation covered in this paper. However, they do have a representation based on expanding periodic covariance functions into a series of stochastic resonators (Appendix~\ref{ch5_sec_periodic}). Thus, the functions $f_2(t)$ and $f_3(t)$ are approximated as in equation (\ref{ch5_eq_f_period}), with variance coefficients $\tilde{q}_j^2$ as in equation (\ref{ch5_eq_q_2}).
We use $J=10$ cosine terms. The lengthscale estimates $\hat{\ell}_2$ and $\hat{\ell}_3$ for the GP components $f_2(t)$ and $f_3(t)$, respectively, are bigger than the minimum lengthscale reported by Figure \ref{ch5_fig28_m_lscale_periodic} as function of the number of cosine terms $J$, which means that the approximations are good.
